### 2018-08-16 Manuscript/Manuscript.Rmd
Small edits

---
title: 'Parsing medial prefrontal cortex: A joint meta-analytic and graph-theoretic approach.'

# Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
author:
  - name: Claudio A. Toro-Serey
    affiliation: a,1,2
  - name: Joseph T. McGuire
    affiliation: a,1
address:
  - code: a
    address: Boston University, Department of Psychological and Brain Sciences, 64 Commonwealth Ave., Boston, 02250

corresponding_author:
  - code: 2
    text: "To whom correspondence should be addressed. E-mail: ctoro@bu.edu"

# For footer text
lead_author_surname: Toro-Serey

## Remove this if not required    
equal_authors:
  - code: 1
    text: "A.O.(Author One) and A.T. (Author Two) contributed equally to this work (remove if not applicable)."

author_contributions: |
  Please provide details of author contributions here.

## Remove this if not required
conflict_of_interest: |
  Please declare any conflict of interest here.

abstract: |
  (Current count: 153) Valuation effects are consistently observed in medial prefrontal and posterior cingulate cortex (mPFC and PCC). The spatial extent of these effects is mostly indistinguishable from the default mode network (DMN) in existing meta-analyses. However, little is known about how valuation effects fit within the broader functional architecture of mPFC and PCC, or whether that architecture is consistent or idiosyncratic across individuals. Here we complement a meta-analysis with fMRI-based graph theoretic approaches to subdivide mPFC and PCC at the single-subject level. Our results suggest the functional topography of mPFC has substantial variability across individuals. This highlights the potential usefulness of estimating brain effects at the individual level in this region, and points to limitations of aggregative methods such as coordinate-based meta-analysis in determining whether valuation and DMN effects emerge from common or separable brain systems. Our approach shows promise in addressing this issue through future manipulations of valuation.

significance: |
  Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality. The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership. The Significance Statement appears in the paper itself and is required for all research papers.

acknowledgements: |
  Please include your acknowledgments here, set in a single paragraph. Please do not include any acknowledgments in the Supporting Information, or anywhere else in the manuscript.

keywords:
  - Networks
  - DMN
  - Valuation

## must be one of: pnasresearcharticle (usual two-column layout), pnasmathematics (one column layout), or pnasinvited (invited submissions only)
pnas_type: pnasresearcharticle

bibliography: pnas-sample.bib
csl: pnas.csl

## change to true to add optional line numbering
lineno: false

output: rticles::pnas_article
---

(current, 676 words. Aim for 650)

Studies of decision making have consistently associated activity in vmPFC with the subjective value assigned to decision outcomes (Bartra, McGuire, & Kable, 2010; Kable & Glimcher, 2007)

The medial prefrontal cortex (mPFC) has been suggested to subserve many of the cognitive abilities that differentiate humans from other animals (Barbas & Garcia-Cabezas, 2016). Specifically, this region has been related to decision making (Bartra, McGuire, & Kable, 2013), memory encoding (Shacter, Addis, & Buckner, 2007), and default mode deactivation (DMN) (Yeo et al., 2011), among ohters. While much has been discovered about this area in primates (Barbas & Garcia-Cabezas, 2016), the lack of direct measurements of neuronal activity makes this a tough area to record in humans, one that nowadays relies mostly on functional MRI (fMRI). Unfortunately, the mPFC is especially challenging to image, as the oxygen-related signal recorded by fMRI is contaminated by the oxygen in the sinuses (Logothetis, 2008). In addition, this area is subject to more idiosyncratic cortical folding than any other region (Zilles, Palomero- Gallagher, & Amunts, 2013), thus adding a level of complexity to the generalizable dissection of topographic functional roles within mPFC.

Previous meta analytic studies have provided important insights on the psychological phenomena attributed to mPFC subregions. For example, de la Vega et al. (2016) found subdivisions of mPFC and MCC that are significantly tied to decision making in the literature. In particular, they show that more ventral regions are more tightly associated with reward and fear. However, this pattern overlaps with their finding that vmPFC also coactivates with DMN regions, and find no topographical distinction between these phenomena. Similarly, the results from Bartra et al and Laird et al show that this region can't be subdivided into DMN vs subjective value (acikalin, chris, russ). However, meta analyses rely on peaks often derived from group-averaged data. Recent work has prescribed relevance to the analysis of single subjects in fMRI. Yeo et al (2018) propose the use of individualized topologies when identifying functional networks. Gordon et al identify important idiosincrasies in oversampled subjects. Similarly, Braga and Buckner show that finer subdivisions of the DMN are possible when looking at each subject independently. This opens up the possibility that distinctions between DMN and subjective value are possible at this finer level of resolution, but would otherwise be masked in meta analytic work.

Braga & Buckner's work is particularly informative when it comes to subdividing DMN. However, the seed-based functional connectivity analysis is sensitive to the selection of the seed. A natural next step would be to recruit all possible patterns of covariation within this network. A popular alternative is to rely on graph theoretic methods, which provide a characterization of networks that organically fits brain dynamics (provides an organic avenue to understand brain dynamics?), and thus permits the consideration of the full covariance structure of brain networks. (ADV & Tal somewhere here).

Given the shortcomings of study-specific and cross-study analyses mentioned above, there is potential value in combining complementary strengths from both approaches. Focusing on resting state data has been successful in determining a number of neural phenomena, including disentangling auditory and visual attention areas (Tobyne, Osher, Michalka, & Somers, 2017).

One great advantage of network science is the ability to characterize connectivity in a way that allows for comparisons among individuals in a meaningful way (Garcia, Ashourvan, Muldoon, Vettel, & Bassett, 2018), thus letting us address the idiosyncrasies mentioned above. In this study, we aim to subdivide mPFC into subject-specific DMN and non-DMN areas, so that we can generate more informed topographic targets for future studies of decision making. We address this aim at 3 levels of granularity: 1) across the literature (meta-analysis), selecting a number of regions of interest (ROI) that show significant overlap and specificity between DMN and decision making; 2) full-brain connectivity of actual fMRI data, using a pre-specified brain atlas; and 3) characterizing the connectivity of surface vertices (small units of function that contain oxygen-dependent neural information across time) for the regions selected in step one. I performed the last two steps at an individual subject level for 3 subjects, hoping to find some commonalities among them. In this way, I attempted to extract functional topographical information from topological network features.

``` {r Libraries and functions, include = FALSE, echo = FALSE}
setwd('../')
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

Cols <- c("aquamarine4","#D9541A",rgb(190,190,190,100, maxColorValue = 255)) # left, right, interhemisphere

# In case I want to visualize stuff in the brain
labelCoords_vertex <- read.csv2('labelCoords_vertex.csv', sep = ",")[,2:6]
labelCoords_vertex <- transform(labelCoords_vertex, x = as.numeric(as.character(x)), y = as.numeric(as.character(y)), z = as.numeric(as.character(z)))

library(igraph)
library(pander)
library(ggnetwork)
library(ggplot2)
library(gridExtra)
library(data.table)
library(plyr)
library(mcclust)
library(lme4)
library(parallel)
library(corrplot)
library(DescTools)

## Cohen's D for 2 groups
# for a more flexible approach, make the data input to be a list with entries for n groups, 
# then do length(list) for the number of groups. 
cohenD <- function(group1 = 1, group2 = 2){

  # means
  mean1 <- mean(group1, na.rm = T)
  mean2 <- mean(group2, na.rm = T)

  # variance
  var1 <- var(group1, na.rm = T)
  var2 <- var(group2, na.rm = T)

  # equation
  out <- (mean1 - mean2) / sqrt((var1 + var2)/2)

  return(out)

}

## log-log degree distribution
logDegreeDist <- function(Graph = padjMat) {

  tempD <- degree(Graph)
  tempDD <- degree.distribution(Graph) # frequency of occurrencies of certain degrees
  d <- (0:(max(tempD)-1))
  ind <- (tempDD!=0)
  plot(d[ind], tempDD[ind], log = "xy", col = "blue", 
       xlab = "Log-Degree", ylab = "Log-Intensity",  
       main = "Log-Log Degree Distribution")

}

## Create correlation matrix between vertices of 2 ROIs
# Requires labeled time series matrix
vertexCorrMat <- function(ROI_1 = 'R_7m_ROI', ROI_2 = 'L_7m_ROI'){

  # extract the time series from the ROIs
  indexing_1 <- which(rownames(timeSeries)==as.character(ROI_1))
  indexing_2 <- which(rownames(timeSeries)==as.character(ROI_2))
  nVerts_1 <- length(indexing_1)
  nVerts_2 <- length(indexing_2)
  ROI_1_tseries <- timeSeries[indexing_1, ]
  ROI_2_tseries <- timeSeries[indexing_2, ]

  # Create empty matrix
  corrMat <- matrix(data = 0, nrow = nVerts_1, ncol = nVerts_2)

  # 
  for (vertex_1 in seq(nVerts_1)){

    # temp series 1
    tempOne <- ROI_1_tseries[vertex_1, ]

    for(vertex_2 in seq(nVerts_2)){

      # temp series 2
      tempTwo <- ROI_2_tseries[vertex_2, ]

      # store the Pearson correlation in the corr matrix
      corrMat[vertex_1, vertex_2] <- cor(tempOne, tempTwo, method = "pearson")

    }
  }

  # name the dimensions of the matrix according to the surface vertex index
  rownames(corrMat) <- indexing_1
  colnames(corrMat) <- indexing_2

  # plot (optional?)  
  # unclustered heatmap
  heatmap(corrMat,
          Rowv = NA,
          Colv = NA,
          scale = "row",
          col = pallette(1000))

  # clustered heatmap
  heatmap(corrMat,
          scale = "row",
          col = pallette(1000))

  # print min-max corr vals
  print(range(corrMat))

  return(corrMat)

}

## Correlation matrix between an ROI and the Glasser parcels
parcelCorrMat <- function(ROI = 'R_7m_ROI', lbels = lookup$V1){

  # extract the time series from the ROIs
  indexing <- which(rownames(timeSeries)==as.character(ROI))
  nVerts <- length(indexing)
  ROI_tseries <- timeSeries[indexing, ]

  # create empty matrix to store values
  corrMat <- matrix(data = 0, nrow = nVerts, ncol = 360)

  # Loop through every combination of label and 7m vertex
  for (Parcel in lbels){

    # Get an ROI index to retrieve timeseries and store correlation vals
    indxROI <- which(lbels==as.character(Parcel))

    for(vertex in seq(nVerts)){

      # temp series
      tempROI <- as.numeric(ptSeries[indxROI, ])
      tempVertex <- ROI_tseries[vertex, ]

      # store the Pearson correlation in the corr matrix
      corrMat[vertex, indxROI] <- cor(tempROI, tempVertex, method = "pearson")

    }
  }

  # name the columns according to labels
  colnames(corrMat) <- lbels
  rownames(corrMat) <- indexing

  # plot (optional?)  
  # unclustered heatmap
  heatmap(corrMat,
          Rowv = NA,
          Colv = NA,
          scale = "none",
          col = pallette(1000))

  # clustered heatmap
  heatmap(corrMat,
          scale = "none",
          col = pallette(1000))

  # print min-max corr vals
  print(range(corrMat))

  return(corrMat)

}

## Correlation matrix between an Parcel and every other grayordinate
# THIS CAN BE OPTIMIZED WITH A BETTER USE OF COR()
partoverCorrMat <- function(ROI = 'R_7m_ROI', lbels = verts[[1]]){

  # extract the time series from the ROI
  indexing <- which(rownames(ptSeries)==as.character(ROI))
  ROI_tseries <- as.numeric(ptSeries[indexing, ])

  # get the remaining number of vertices
  nVerts <- length(lbels)

  # Create data frame to store values
  parcelIndx <- which(labelCoords_parcel$Label==ROI)
  corrVec <- data.frame(Seed = rep(ROI, nVerts),
                        Labels = lbels,
                        xstart = rep(labelCoords_parcel$x[parcelIndx], nVerts), # 7.33324 for R_7m
                        ystart = rep(labelCoords_parcel$y[parcelIndx], nVerts), # -63.142
                        zstart = rep(labelCoords_parcel$z[parcelIndx], nVerts)) # 42.5578

  for(vertex in seq(nVerts)){

    # temp series
    tempVertex <- timeSeries[vertex, ]

    # store the Pearson correlation in the corr matrix
    #corrVec$Correlation[vertex] <- cor(ROI_tseries, tempVertex, method = "pearson")

    # Another option, so that FDR can be applied
    tempCor <- cor.test(ROI_tseries, tempVertex, method = "pearson")
    corrVec$Correlation[vertex] <- tempCor$estimate
    corrVec$pval[vertex] <- tempCor$p.value

  }

  # Turn ROI vertices into 0s
  indexing <- which(corrVec$Labels==as.character(ROI))
  corrVec$Correlation[indexing] <- 0

  # Create a column with p-vals corrected for multiple comparisons
  corrVec$adjPval <- p.adjust(corrVec$pval, "BY")

  # normalize correlation vals for analysis (see below)
  tempTanh <- fisherTanh(Data = corrVec$Correlation)
  corrVec$tanhZ <- tempTanh$tanhZ
  corrVec$tanhPvals <- tempTanh$pvals
  corrVec$tanhPAdjusted <- tempTanh$adjustPvals

  return(corrVec)

}

## tanh-z transformation (variance stabilizing Fisher) and p-values (adjusted and not)
# This takes either a matrix of correlation values (vectors too, but manually compute pvals)
# Normalization approach suggested in network textbook (equation 7.10)
fisherTanh <- function(Data = padjMatrix){

  transformed <- list()

  # tanh
  transformed$tanhZ <- 0.5 * log((1 + Data) / (1 - Data))

  # p-vals
  if (is.matrix(Data)) {
    z.vec <- transformed$tanhZ[upper.tri(transformed$tanhZ)]
    n <- dim(Data)[1]
  } else if (is.vector(Data)) {
    z.vec <- transformed$tanhZ
    n <- length(Data)
  }
  transformed$pvals <- 2 * pnorm(abs(z.vec), 0 , sqrt(1 / (n-3)), lower.tail=F)

  # adjust pvals
  transformed$adjustPvals <- p.adjust(transformed$pvals, "BH")

  if (is.matrix(Data)) {
    # get pvals and their adjustment into a symetric matrix form
    # regular
    tempMat <- matrix(0, dim(Data)[1], dim(Data)[2])
    tempMat[upper.tri(tempMat)] <- transformed$pvals
    tempMat[lower.tri(tempMat)] <- transformed$pvals
    dimnames(tempMat) <- list(rownames(Data), rownames(Data))
    transformed$pvals <- tempMat

    # adjusted
    tempMat <- matrix(0, dim(Data)[1], dim(Data)[2])
    tempMat[upper.tri(tempMat)] <- transformed$adjustPvals
    tempMat[lower.tri(tempMat)] <- transformed$adjustPvals
    dimnames(tempMat) <- list(rownames(Data), rownames(Data))
    transformed$adjustPvals <- tempMat
  }

  return(transformed)

}

## Extract the centroid-most vertex from each parcel
parcelCentroid <- function(ROI = 'R_7m_ROI', all_coordinates = labelCoords_vertex){

  # The coordinate file must contain a Label column

  # Get the ROI-specific vertices
  indx <- grep(ROI, all_coordinates$Label)

  # Grab the coordinates
  tempCoords <- all_coordinates[indx, c("x","y","z")]
  rownames(tempCoords) <- indx

  # Compute the distance among all vertices, and turn into a matrix
  tempDist <- dist(tempCoords, diag = T)
  tempDist <- as.matrix(tempDist)

  # Compute the sum of distances for each vertex, and get the minimum
  sumDists <- colSums(tempDist)
  minDist <- which(sumDists == min(sumDists))
  minDist <- indx[minDist]

  return(minDist)

}

## This function takes the output from partoverCorrMat() and adds/removes the characteristics we decided on.
prepROItoVer <- function(Data = R_7m_allCorr, Coordinates = labelCoords_vertex){

  # Add vertex coordinates
  tempDframe <- cbind(Data, Coordinates[,1:3])

  # Remove rows with non-significant adjusted pvalues
  indx <- tempDframe$adjPval < 0.05
  tempDframe <- tempDframe[indx, ]

  # Characterize correlation direction and round up vals for plotting
  tempDframe$Relation <- ifelse(tempDframe$Correlation < 0, "Negative", "Positive")
  tempDframe$Correlation <- round(tempDframe$Correlation, digits = 2)

  # Add a column to differentiate between left/right hemispheres
  tempDframe$Hemisphere <- substring(tempDframe$Labels,1,1)

  return(tempDframe)

}

## Plot correlation from ROI to rest of vertices
plotROItoVertex <- function(Data = R_7m_allCorr, ROI = 'R_7m_ROI', ColRange = Cols, View = "Axial", Legends = TRUE){

  ROIindx <- grep(ROI, Data$Labels)
  ROIvertices <- Data[ROIindx,c("x","y","z")] # grab ROI-specific vertices to black out
  labelCoord <- ROIvertices[1,] # just to place the label

  if (View == "Axial") {
    ggplot() +
      geom_point(data = Data, aes(x=x, y=y, alpha=.5, color = Correlation), show.legend = Legends) +
      geom_point(data = labelCoords_vertex, aes(x=x, y=y), alpha=0.01) +
      geom_nodes(data = ROIvertices, aes(x=x, y=y)) +
      geom_nodelabel_repel(aes(x=labelCoord$x, y=labelCoord$y, label = sub("_ROI","",ROI))) +
      scale_color_gradient2(low = ColRange[1], mid = "white", high = ColRange[2], limits = c(-1,1)) +
      theme_blank()
  } else if (View == "Sagittal") {
    ggplot() +
      geom_point(data = Data, aes(x=y, y=z, alpha=.5, color = Correlation), show.legend = Legends) +
      geom_point(data = labelCoords_vertex, aes(x=y, y=z), alpha=0.01) +
      geom_nodes(data = ROIvertices, aes(x=y, y=z)) +
      geom_nodelabel_repel(aes(x=labelCoord$y, y=labelCoord$z, label = sub("_ROI","",ROI))) +
      scale_color_gradient2(low = ColRange[1], mid = "white", high = ColRange[2], limits = c(-1,1)) +
      theme_blank()
  } else if (View == "Coronal") {
    ggplot() +
      geom_point(data = Data, aes(x=x, y=z, alpha=.5, color = Correlation), show.legend = Legends) +
      geom_point(data = labelCoords_vertex, aes(x=x, y=z),  alpha=0.01) +
      geom_nodes(data = ROIvertices, aes(x=x, y=z), alpha=.5) +
      geom_nodelabel_repel(aes(x=labelCoord$x, y=labelCoord$z, label = sub("_ROI","",ROI))) +
      scale_color_gradient2(low = ColRange[1], mid = "white", high = ColRange[2], limits = c(-1,1)) +
      theme_blank()
  } else if (View == "Medial Right") {
    if (substring(ROI,1,1) == "R") {
      oneHemi <- Data[grep("R", Data$Hemisphere), ]
      oneHemi$Medial <- ifelse(oneHemi$x < 20, "Medial", "Other")
      medialHemi <- oneHemi[which(oneHemi$Medial=="Medial"), ]
      medialHemi$y <- medialHemi$y * -1
      ROIvertices$y <- ROIvertices$y * -1
      labelCoord$y <- labelCoord$y * -1
      oneHemi_all <- labelCoords_vertex[grep("R",labelCoords_vertex$Hemisphere), ] # whole surface
      oneHemi_all$Medial <- ifelse(oneHemi_all$x < 20, "Medial", "Other")
      medialHemi_all <- oneHemi_all[grep("Medial", oneHemi_all$Medial), ]
      medialHemi_all$y <- medialHemi_all$y * -1
      ggplot() +
        geom_point(data = medialHemi, aes(x=y, y=z, alpha=.1, color = Correlation), show.legend = Legends) +
        #geom_point(data = medialHemi_all, aes(x=y, y=z), alpha=0.07) +
        scale_color_gradient2(low = ColRange[1], mid = "white", high = ColRange[2], limits = c(-1,1)) +
        geom_nodes(data = ROIvertices, aes(x=y, y=z), alpha=.5) +
        geom_nodelabel_repel(aes(x=labelCoord$y, y=labelCoord$z, label = sub("_ROI","",ROI))) +
        theme_blank()
    } else {
      oneHemi <- Data[grep("R", Data$Hemisphere), ]
      oneHemi$Medial <- ifelse(oneHemi$x < 20, "Medial", "Other")
      medialHemi <- oneHemi[which(oneHemi$Medial=="Medial"), ]
      medialHemi$y <- medialHemi$y * -1
      ROIvertices$y <- ROIvertices$y * -1
      labelCoord$y <- labelCoord$y * -1
      oneHemi_all <- labelCoords_vertex[grep("R",labelCoords_vertex$Hemisphere), ] # whole surface
      oneHemi_all$Medial <- ifelse(oneHemi_all$x < 20, "Medial", "Other")
      medialHemi_all <- oneHemi_all[grep("Medial", oneHemi_all$Medial), ]
      medialHemi_all$y <- medialHemi_all$y * -1
      ggplot() +
        geom_point(data = medialHemi, aes(x=y, y=z, alpha=.1, color = Correlation), show.legend = Legends) +
        #geom_point(data = medialHemi_all, aes(x=y, y=z), alpha=0.07) +
        scale_color_gradient2(low = ColRange[1], mid = "white", high = ColRange[2], limits = c(-1,1)) +
        theme_blank()
    }
  } else if (View == "Medial Left") {
    if (substring(ROI,1,1) == "L") {
      oneHemi <- Data[grep("L", Data$Hemisphere), ]
      oneHemi$Medial <- ifelse(oneHemi$x > -20, "Medial", "Other")
      medialHemi <- oneHemi[which(oneHemi$Medial=="Medial"), ]
      oneHemi_all <- labelCoords_vertex[grep("R",labelCoords_vertex$Hemisphere), ] # whole surface
      oneHemi_all$Medial <- ifelse(oneHemi_all$x < 20, "Medial", "Other")
      medialHemi_all <- oneHemi_all[grep("Medial", oneHemi_all$Medial), ]
      ggplot() +
        geom_point(data = medialHemi, aes(x=y, y=z, alpha=.1, color = Correlation), show.legend = Legends) +
        #geom_point(data = medialHemi_all, aes(x=y, y=z), alpha=0.07) +
        scale_color_gradient2(low = ColRange[1], mid = "white", high = ColRange[2], limits = c(-1,1)) +
        geom_nodes(data = ROIvertices, aes(x=y, y=z), alpha=.5) +
        geom_nodelabel_repel(aes(x=labelCoord$y, y=labelCoord$z, label = sub("_ROI","",ROI))) +
        theme_blank()
    } else {
      oneHemi <- Data[grep("L", Data$Hemisphere), ]
      oneHemi$Medial <- ifelse(oneHemi$x > -20, "Medial", "Other")
      medialHemi <- oneHemi[which(oneHemi$Medial=="Medial"), ]
      oneHemi_all <- labelCoords_vertex[grep("R",labelCoords_vertex$Hemisphere), ] # whole surface
      oneHemi_all$Medial <- ifelse(oneHemi_all$x < 20, "Medial", "Other")
      medialHemi_all <- oneHemi_all[grep("Medial", oneHemi_all$Medial), ]
      ggplot() +
        geom_point(data = medialHemi, aes(x=y, y=z, alpha=.1, color = Correlation), show.legend = Legends) +
        #geom_point(data = medialHemi_all, aes(x=y, y=z), alpha=0.07) +
        scale_color_gradient2(low = ColRange[1], mid = "white", high = ColRange[2], limits = c(-1,1)) +
        theme_blank()
    }
  }

}

## Run a fastgreedy modularity community detection on ROIs
# This function relies on having the timeSeries data uploaded, and labelCoords_vertex 
# Extras dictates whether the community object + correlation matrix should also be extracted
communityDetection <- function(Data = parcelBins$First, ROIS = "None", Type = "vertex", thresh = F, extras = F) {

  print(paste('Computing modularity based on', Type))

  if (ROIS == "None"){

    # This will just do SP for now, for sliding window
    print("Previously concatenated data")

    corrMat <- cor(t(Data))
    corrMatrix <- corrMat
    transfMat <- fisherTanh(Data = corrMat)
    if (thresh == T) {
      transfMat$tanhZ[transfMat$adjustPvals > 0.05] <- 0
    }
    corrMat <- transfMat$tanhZ

    diag(transfMat$tanhZ) <- 0    
    diag(corrMat) <- 0

    corrMat <- exp(corrMat)
    corrMat[corrMat==1] <- 0
    transfMat$tanhZ <- corrMat

    tempGraph <- graph_from_adjacency_matrix(corrMat, weighted = T, mode = "undirected")
    tempLap <- laplacian_matrix(tempGraph, normalized=T)
    tempEigen <- eigen(tempLap)
    f.vec <- length(tempEigen$values) - 1
    tempEigen$binarized <-  as.factor(ifelse(tempEigen$vectors[,f.vec] > 0, 1, 0)) # binarized Fiedler Vector

    summary <- data.frame(Label = colnames(corrMat),
                          Hemisphere = substring(colnames(corrMat),1,1),
                          EigenVal = tempEigen$values,
                          FiedlerVec = tempEigen$vectors[, (length(tempEigen$values) - 1)],
                          FiedlerBinary = tempEigen$binarized)

    ## Get the final components
    if (extras == T) {
      modularityResults <- list(CorrMatrix = corrMatrix,
                                TransfMatrix = transfMat,
                                Summary = summary)
    } else {
      modularityResults <- list(Summary = summary)
    }

  } else {

    if (Type == "vertex") {
      # To store the vertex indices corresponding to the ROIs
      indx <- numeric()

      # had to place dashes on each side because grep grabbed strings containing the names (i.e. 47m, a24pr)
      #ROIS <- c("_7m_", vmPFC_labels) # c("_a24_", "_7m_")
      for (ROI in ROIS) {
        indx <- c(indx, grep(ROI, rownames(Data)))  
      }

      nVerts <- length(indx)
      ROI_tseries <- Data[indx, ]

      # This used to be done with the for loop, but it was too slow. cor() speeds up the process by a lot
      corrMat <- cor(t(ROI_tseries))

      # name the dimensions of the matrix according to the surface vertex index
      rownames(corrMat) <- indx

      # for storing later
      corrMatrix <- corrMat

      # transform to Fisher's (think of thresholding)
      transfMat <- fisherTanh(Data = corrMat)

      # Determine if edges should be thresholded or not
      if (thresh == T) {
        transfMat$tanhZ[transfMat$adjustPvals > 0.05] <- 0
      }

      # Store Fisher transformed vals for graphing
      corrMat <- transfMat$tanhZ

      # diagonals of 1 could be interpreted as self-loops
      diag(transfMat$tanhZ) <- 0    
      diag(corrMat) <- 0

      # Exponentiate to preserve distribution while ensuring positive weights
      # I'm keeping corrMat and transfMat$tanhZ separate in case I want to uncorrect transfMat in the future
      corrMat <- exp(corrMat)
      corrMat[corrMat==1] <- 0
      transfMat$tanhZ <- corrMat

      # community detection
      # I initially used the absolute value of the correlation, but the exp preserves the distribution 
      # Next, try using the fisher transform

      tempGraph <- graph_from_adjacency_matrix(corrMat, weighted = T, mode = "undirected")
      tempCommunity <- fastgreedy.community(tempGraph)

      # community object
      #modularityResults$community <- tempCommunity

      # correlation matrix with transformed values
      #modularityResults$corrMat <- corrMatrix

      # get coordinate info from selected regions (useful for ggplot)
      summary <- data.frame(Vertex = indx,
                            Label = tempCommunity$names,
                            Membership = tempCommunity$membership,
                            Modularity = tempCommunity$modularity,
                            x = labelCoords_vertex[indx, "x"],
                            y = labelCoords_vertex[indx, "y"],
                            z = labelCoords_vertex[indx, "z"],
                            Hemisphere = substring(tempCommunity$names,1,1))

    } else if (Type == "parcels") {

      # This used to be done with the for loop, but it was too slow. cor() speeds up the process by a lot
      corrMat <- cor(t(Data))

      # for storing later
      corrMatrix <- corrMat

      # transform to Fisher's (think of thresholding)
      transfMat <- fisherTanh(Data = corrMat)

      # Determine if edges should be thresholded or not
      if (thresh == T) {
        transfMat$tanhZ[transfMat$adjustPvals > 0.05] <- 0
      }

      # Store Fisher transformed vals for graphing
      corrMat <- transfMat$tanhZ

      # diagonals of 1 could be interpreted as self-loops
      diag(transfMat$tanhZ) <- 0
      diag(corrMat) <- 0

      # Exponentiate to preserve distribution while ensuring positive weights
      # I'm keeping corrMat and transfMat$tanhZ separate in case I want to uncorrect transfMat in the future
      corrMat <- exp(corrMat)
      corrMat[corrMat==1] <- 0
      transfMat$tanhZ <- corrMat

      # community detection
      # I initially used the absolute value of the correlation, but the exp preserves the distribution 
      # Next, try thresholding it by the adjusted p-vals
      tempGraph <- graph_from_adjacency_matrix(corrMat, weighted = T, mode = "undirected")
      tempCommunity <- fastgreedy.community(tempGraph)

      # community object
      #modularityResults$community <- tempCommunity

      # correlation matrix with transformed values
      #modularityResults$corrMat <- corrMatrix

      # get coordinate info from selected regions (useful for ggplot)
      summary <- data.frame(Label = tempCommunity$names,
                            Membership = tempCommunity$membership,
                            Modularity = tempCommunity$modularity,
                            x = labelCoords_parcel[ ,"x"],
                            y = labelCoords_parcel[ ,"y"],
                            z = labelCoords_parcel[ ,"z"],
                            Hemisphere = substring(tempCommunity$names,1,1))

    }

    ## Get the final components
    modularityResults <- list(Community = tempCommunity,
                              CorrMatrix = corrMatrix,
                              TransfMatrix = transfMat,
                              Summary = summary)

  }

  return(modularityResults)

}

## plot the communities from communityDetection
plotCommunities <- function(Data = modularityResults, Hemi = "R", type = "Membership", Legends = T, bground=0.1, Cols = c("aquamarine4", "#D9541A")) {
  # This takes the output from the communityDetection function and plots them on the medial wall (since we're interested in medial now)
  # type is the name of the column to plot
  if (Hemi == "R") {
    # "Medial Right"
    oneHemi <- Data[grep("R", Data$Hemisphere), ] # from the ROIs
    oneHemi$Medial <- ifelse(oneHemi$x < 20, "Medial", "Other")
    medialHemi <- oneHemi[grep("Medial", oneHemi$Medial), ]
    medialHemi$y <- medialHemi$y * -1
    oneHemi_all <- labelCoords_vertex[grep("R",labelCoords_vertex$Hemisphere), ] # whole surface
    oneHemi_all$Medial <- ifelse(oneHemi_all$x < 20, "Medial", "Other")
    medialHemi_all <- oneHemi_all[grep("Medial", oneHemi_all$Medial), ]
    medialHemi_all$y <- medialHemi_all$y * -1
    ggplot() + 
      geom_point(data = medialHemi, aes_string(x="y", y="z", color = type), show.legend=Legends) +
      geom_point(data = medialHemi_all, aes(x=y, y=z), alpha=bground) +
      scale_color_gradient(low = Cols[1], high = Cols[2]) +
      theme_blank() 
  } else if (Hemi == "L") {
    # "Medial Left"
    oneHemi <- Data[grep("L", Data$Hemisphere), ]
    oneHemi$Medial <- ifelse(oneHemi$x > -20, "Medial", "Other")
    medialHemi <- oneHemi[which(oneHemi$Medial=="Medial"), ]
    oneHemi_all <- labelCoords_vertex[grep("L",labelCoords_vertex$Hemisphere), ] # whole surface
    oneHemi_all$Medial <- ifelse(oneHemi_all$x > -20, "Medial", "Other")
    medialHemi_all <- oneHemi_all[grep("Medial", oneHemi_all$Medial), ]
    ggplot() + 
      geom_point(data = medialHemi, aes_string(x="y", y="z", color = type), show.legend=Legends) +
      geom_point(data = medialHemi_all, aes(x=y, y=z), alpha=bground) +
      scale_color_gradient(low = Cols[1], high = Cols[2]) +
      theme_blank() 
  }
}

## get ROI coords & index
# The point here is to reduce the summary dframes from community detection to show only ROIs
# Should work for extracting any label-indexed dframe though
# I wanted to also get the index in case I want to extract specific rows from parcel/vertex coord dframes
getCoords <- function(Labels = DMN_labels, Coords = labelCoords_parcel, TimeSeries = FALSE){

  indx <- numeric()

  # If you want to select time series from raw data
  if (TimeSeries == TRUE) {

    for (ROI in Labels) {
      indx <- c(indx, grep(ROI, rownames(Coords)))
    }

  } else { # for the summary output of the community detection output

    for (ROI in Labels) {
      indx <- c(indx, grep(ROI, Coords$Label))
    }

  }

  results <- list()
  results$Index <- indx
  results$Coords <- Coords[indx, ]

  return(results)

}

# eigen value community detection
eigenVals <- function(Data = binnedCommunities_p$First) {

  # This will produce a vector of values, ready to be plotted
  tempGraph <- graph_from_adjacency_matrix(Data$TransfMatrix$tanhZ, mode = "undirected", weighted = T)
  tempGraph <- laplacian_matrix(tempGraph, normalized=T)
  tempGraph <- eigen(tempGraph)
  f.vec <- length(tempGraph$values) - 1
  tempGraph$binarized <-  as.factor(ifelse(tempGraph$vectors[,f.vec] > 0, 1, 0)) # binarized Fiedler Vector

  return(tempGraph)

}

# Attempt at setting up data for confusion matrices and Jaccard index calculations
confusionMatrix <- function(partition1 = parcelCommunities[[1]]$Membership, partition2 = parcelCommunities[[1]]$FiedlerBinary) {

  # Divide partitions to evaluate
  # Usually 2 will be fiedler

  # Check if any partition has 0s (since I binarize the Fiedler vector)
  if (0 %in% partition1) {

    partition1[grep(0, partition1)] <- 1
    partition1[grep(1, partition1)] <- 2

  }

  if (0 %in% partition2) {

    partition2[grep(1, partition2)] <- 2   
    partition2[grep(0, partition2)] <- 1

  }

  # Community sizes
  commSizes1 <- table(partition1)
  commSizes2 <- table(partition2)

  # Number of communities per partition
  nComms1 <- length(commSizes1)
  nComms2 <- length(commSizes2)

  # Number of vertices
  n <- length(partition1)

  # Putting together elements of the confusion matrix
  confMatrix <- matrix(nrow = nComms1,
                       ncol = nComms2)

  for (i in as.numeric(case.names(commSizes1))) {

    # Vertices belonging to community qX of partition X
    tempComm1 <- partition1 == i

    for (j in as.numeric(case.names(commSizes2))) {

      # Vertices belonging to community qY of partition Y
      tempComm2 <- partition2 == j

      # Populate matrix
      confMatrix[i,j] <- sum(tempComm1 & tempComm2)

    }
  }

  if (sum(confMatrix) != n) {warning('Sum of the confusion matrix is not equal to number of vertices')}

  return(confMatrix)

}

# Permutation for 2 groups
permute <- function(group1 = 1, group2 = 2, statType = mean, nPerms = 5000, paired = FALSE){

  # prep data
  summaryPerm <- list()
  lOne <- length(group1)
  lTwo <- length(group2)
  bigSample <- c(group1,group2)

  if (paired == FALSE) {

    for (i in 1:nPerms){

      # relabel samples
      tempBig <- sample(bigSample)
      tempOne <- tempBig[1:lOne]
      tempTwo <- tempBig[(lOne+1):length(bigSample)]

      # stats
      tempDiffs <- statType(tempOne,na.rm=T) - statType(tempTwo,na.rm=T)
      summaryPerm$jointDist[i] <- tempDiffs # statType(tempDiffs, na.rm = T)

    }

  } else {

    for (i in 1:nPerms){

      # shift labels in a pairwise fashion
      tempDiffs <- statType((-1)^rbinom(lOne,1,0.5) * (group1 - group2))
      summaryPerm$jointDist[i] <- tempDiffs

    }

  }

  # get the observed difference
  diffs <- statType(group1,na.rm=T) - statType(group2,na.rm=T)
  observedAbs <- abs(diffs) # maybe leave it as means here
  observed <- diffs
  summaryPerm$Pval <- 2 * (1 - ecdf(summaryPerm$jointDist)(observedAbs))
  if (length(unique(abs(summaryPerm$jointDist))) == 1) {summaryPerm$Pval <- 1} # if the difference is always the same, then p = 1
  summaryPerm$Observed <- observed

  return(summaryPerm)

}

# Non-parametric Bootstrap for a single group
bootstrap <- function(group = 1, statType = mean, B = 5000){

  # prep param
  bootStats <- rep(0,B)

  # iterate
  for(b in 1:B){

    # wait group
    x <- sample(group,length(group),replace=T)  
    bootStats[b] <- statType(x,na.rm = T)

  }

  return(bootStats)

}

# number of times a node changes affiliation in a time series
flexibility <- function(Data = cbind(1:10, 6:15)) {
  # Calculation of flexibility based on Garcia et al., 2018  
  # This function takes in a data frame or matrix in which columns are the community affiliation 
  # It needs at least 2 colums, although it's pointless for that

  # How many jumps can there be?
  nJumps <- dim(Data)[2] - 1

  # Matrix to store jumps
  jumpCount <- matrix(nrow = dim(Data)[1], ncol = nJumps)

  # For each transition, get the which nodes changed affiliation
  for (jump in seq(nJumps)) {

    jumpCount[, jump] <- Data[, jump] != Data[, jump + 1]

  }

  # How many times did every node jump?
  totalJumps <- rowSums(jumpCount)

  # Calculate the flexibility per node
  flexibility <- totalJumps / nJumps

  return(flexibility)

}

# Sliding window
slidingWindow <- function(subjTS = ROI_timeSeries[[1]], mins = 15, jump = 1, Spectral = T, Modularity = F, ROIs = c("_7m_", vmPFC_labels)) {
  # This is a fairly specific function. It takes the time series from a participant and preps/runs community detection at each specified time window. 
  # Returns the summaries for each window (based on communityDetection function)
  # The selection of a window size is based on the fact that a TR = 0.720s and an hour is 5000 TRs (HCP-based)
  # Since each subject has a different amount of time points, and all are slightly under 1 hr, I chose to round down the number of window moves to avoid unevenness
  # 
  # Parameter definitions:
  #   
  #   subjTS: Subject time series. Note: if you want to look at more than vmpfc and 7m, add a parameter for the labels that can be passed to the community detection function
  # 
  #   mins: Size of the window that will slide through the data.
  #   
  #   jump: Steps (in mins) advanced per slide
  #   
  #   Spectral: Whether to compute spectral partitioning as well. Takes significantly longer, but might be more useful for bisections
  #
  # Right now this takes ~27 mins per subject to run. Think of ways to improve that.

  # Adapted so it works with the output from getCoords
  # One wouldn't really apply a full-brain analysis of this sort anyways, too computationally intensive
  indx <- subjTS$Index
  subjTS <- subjTS$Coords

  # Window sizing (length)
  # Think about incorporating custom TRs
  TS <- dim(subjTS)[2] # time series for the subject
  WS <- seq(834 * (mins/10)) # window from the first TR up to 834 (~10 mins) times the desired multiplier
  jump <- 84 * jump # 84 ~ 1 min, times the number of mins that the window moves
  nJumps <- floor((TS - length(WS)) / jump) # number of jumps to be performed, based on the selected parameters

  ##------- using lapply
  winData <- mclapply(seq(nJumps), function(x) subjTS[,WS+(jump*(x-1))])
  commTS <- mclapply(winData, communityDetection, ROIS = "None", Type = "vertex", thresh = T, extras = F)
  return(commTS)

}

# compare sliding window data to overall communities
slideCompare <- function(subjData = slideCommunities[[1]], template = vmpfc7mCommunities[[1]], func = "RI", comm = "Spectral") {
  # This function compares the community partition from each window slide to the one derived from the whole data set
  # 
  # Parameter definitions:
  #   
  #   subjData: A participant's output from slideCommunities
  #   
  #   template: The partition from the whole time series
  # 
  #   func: Which function to use for comparing ("RI" for adjusted RI, "VI" for variation of information)
  # 
  #   comm: The subject data might contain modularity and spectral partitions. Choose which to use.

  # For storing RIs or VIs
  tempComparison <- numeric()

  # How many jumps does the original data contain?
  nJumps <- length(subjData)

  # This loop is technically backwards. I should technically divide by function, then partition method, and then run the window comparisons
  # It's still really fast, so I won't worry.
  for (Win in seq(nJumps)) {
    if (comm == "Modularity") {
      if (func == "RI") {
        tempComparison[Win] <- arandi(subjData[[Win]]$Membership, template$Membership, adjust = T)
      } else if (func == "VI") {
        tempComparison[Win] <- vi.dist(subjData[[Win]]$Membership, template$Membership)
      }
    } else if (comm == "Spectral") {
      if (func == "RI") {
        tempComparison[Win] <- arandi(subjData[[Win]]$FiedlerBinary, template$FiedlerBinary, adjust = T)
      } else if (func == "VI") {
        tempComparison[Win] <- vi.dist(subjData[[Win]]$FiedlerBinary, template$FiedlerBinary)
      }
    }
  }

  return(tempComparison)

}

# Ensure that all spectral communities associated with 7m (i.e. DMN) have the same label value of 1 
evenSpectral <- function(Data = slideCommunities[[1]][[7]]) {

  # Get only the values for 7m
  shortData <- Data[grep("_7m_", Data$Label), ]

  # Get 7m's most probable affiliation
  affil <- mean(as.numeric(shortData$FiedlerBinary)-1)

  # If it is close to 1, then invert the labeling. 
  # Note: This has no effect on RI or VI, since they are insensitive to actual labeling
  # This is for visualization purposes only
  if (affil < 0.5) {

    UD <- as.numeric(Data$FiedlerBinary) - 1
    Data$FiedlerBinary <- as.factor((UD - 1)^2)
    Data$FiedlerVec <- Data$FiedlerVec * -1

  }

  return(Data)

}

# Create a vector ready to be used for HCP data (32k CIFTI surface)
# The input should be the summary from community partitioning
# Once this is created, go to the terminal and input something like this
# wb_command -cifti-convert -from-text dataforCifti.txt 100307.MyelinMap_BC.32k_fs_LR.dscalar.nii testCifti.dscalar.nii
# Where the myelin file here is just a templace. It can be any dscalar.nii with the right surface size
HCPOut <- function(Data = dmnval7mCommunities[[1]], MOI = "Membership", SubjID = "100307"){

  nVertices <- 59412
  tempVec <- rep(-1, nVertices)
  temp <- grep(MOI, colnames(Data))
  tempVec[Data$Vertex] <- Data[[temp]]
  write.table(file = paste(SubjID,"_",MOI,'_dataforCifti.txt', sep=""), tempVec, row.names = F, col.names = F, dec = ".")

}

# Perform pairwise comparisons of clustering outcomes on all subjects
comparePartitions <- function(Data = dmnval7mCommunities, MOI = "FiedlerBinary", Index = "VI", nSubjects = nSubj, subjNames = subjList) {
  # This function will compare the community partitions from all subjects and create a 'comparison matrix' for every pairwise combination of subjects
  # Alternatively, if a second MOI is added 
  # Inputs
  # 
  # Data: the list of summaries produced by the script
  # 
  # MOI: measure of interest (usually the binarized Fiedler vector). If a vector, compares across algorithms per subject
  # 
  # Index: VI for variation of information, RI for the adjusted rand index, Cor for a Pearson correlation

  # Get the column position of the MOI(s)
  Columns <- colnames(Data[[1]])
  indx <- as.numeric(Columns %in% MOI)
  MOI_indx <- which(indx==1)

  if (length(MOI_indx) < 2) {
    # Combine the measures of interest
    allVecs <- do.call(cbind, lapply(Data, "[[", MOI))

    # Create empty matrix
    indexMatrix <- matrix(data = NA, nrow = nSubjects, ncol = nSubjects)
    dimnames(indexMatrix) <- list(subjNames, subjNames)

    # Run every pairwise comparison with the index of interest on the measure of interest
    if (Index == "VI") {
      for (subj in seq(nSubjects)) {
        for (subj2 in seq(nSubjects)) {
          indexMatrix[subj,subj2] <- vi.dist(allVecs[, subj], allVecs[, subj2])
        }
      }
    } else if (Index == "RI") {
      for (subj in seq(nSubjects)) {
        for (subj2 in seq(nSubjects)) {
          indexMatrix[subj,subj2] <- arandi(allVecs[, subj], allVecs[, subj2], adjust = T)
        }
      }
    } else if (Index == "Cor") {
      for (subj in seq(nSubjects)) {
        for (subj2 in seq(nSubjects)) {
          indexMatrix[subj,subj2] <- cor(allVecs[, subj], allVecs[, subj2])
        }
      }
    }
  } else {
    indexMatrix <- data.frame(SubjID = as.character(subjList),
                              Index = rep(0, nSubjects))
    # Run every pairwise comparison with the index of interest on the measure of interest
    if (Index == "VI") {
      for (subj in seq(nSubjects)) {
        indexMatrix$Index[subj] <- vi.dist(Data[[subj]][, MOI_indx[1]], Data[[subj]][, MOI_indx[2]])
      }
    } else if (Index == "RI") {
      for (subj in seq(nSubjects)) {
        indexMatrix$Index[subj] <- arandi(Data[[subj]][, MOI_indx[1]], Data[[subj]][, MOI_indx[2]], adjust = T)
      }
    } else if (Index == "Cor") {
      for (subj in seq(nSubjects)) {
        indexMatrix$Index[subj] <- cor(Data[[subj]][, MOI_indx[1]], Data[[subj]][, MOI_indx[2]])
      }
    }
  }

  return(indexMatrix)

}

# Generate a matrix with submatrices of 1s of a given size along the diagonal
# Useful to extract specific portions of data from a similarity matrix
diagBlocks <- function(dims = 20, sqSize = 4) {
  div <- dims %/% sqSize
  finalMatrix<-matrix(0,nrow=dims,ncol=dims)
  index <- 1
  for(k in seq(div)){
    finalMatrix[index:(index + (sqSize-1)),index:(index + (sqSize-1))] <- 1
    index<-index + sqSize
  }
  finalMatrix <- finalMatrix == 1
  finalMatrix
}

```

Results {.unnumbered}
========================================
``` {r Partition comparison: all brain example, fig.cap = "Template brain.", fig.width = 4, fig.height = 4, echo = FALSE}
# Load data
setwd('../Summary/')
temp <- list.files()
subjList <- sapply(temp, substring, first=1, last=6)
Summaries <- lapply(temp, read.csv)
nSubj <- length(Summaries)

plotCommunities(Summaries[[1]], Hemi = "L", Legends = F, type = "FiedlerBinary")

```

Discussion {.unnumbered}
========================================

(1500 words max)

Methods {.unnumbered}
========================================

See http://www.jneurosci.org/content/preparing-manuscript

**Meta-analysis**
To perform the first step, I used data from a metanalysis that gathered peak brain coordinates of activity from studies of valuation (n = 27 studies) and DMN (n = 77). These are the surviving areas post-statistical thresholding from each study. I analyzed this data by mapping the peak cortical coordinates to an atlas of the human cortical surface (Glasser et al., 2016). This produced a list of standardized parcels that were reported on each study. The advantage of utilizing this atlas was threefold: first, it reduced the number of functional vertices from 60,000 (cortical surface vertices) to 360 (cortical parcels). Second, it standardizes the vertices from the first to the second stage of my study. Finally, it allowed me to project each area on a brain space without much clutter. Since each vertex is embedded with topographical information, and since functional topography is essential to my question, I mostly project my results onto brain space.

I thus generated two undirected, loop-less, weighted graphs in which each vertex was a brain parcel, and edges correspond to the number of times each pair of them co-occurred across studies. Since different decision tasks use different sensory modalities, and since sensory areas get sometimes reported as unrelated peak coordinates in DMN studies, I bootstrapped the co- occurrences across studies and removed edges whose co-occurrences lied below the 90th percentile. Based on this, only edges with co-occurrences higher than 8 were preserved, and the remaining unconnected vertices (brain regions) were discarded. With these graphs set up, I computed the strength and betweenness centrality in order to capture each parcel’s centrality in the valuation and DMN literatures separately.

In order to select of regions of interest (ROI) from mPFC and other brain areas that are central to valuation and DMN literatures, I first bootstrapped the mean strength for each graph separately, and chose areas with strength above the 95th percentile of this distribution. Next, I took the intersection of these areas, such that the final ROIs were brain regions that overlapped between DMN and valuation in the literature (and thus my main target for dissociation).

**Whole Brain Parcellated connectivity**
For the second step, in order to actually quantify the intrinsic connectivity of these ROIs in the brain, I used data from the Human Connectome Project. Specifically, I acquired resting-state fMRI activity from 3 subjects, all of whom have been preprocessed according to industry standards to allow for group-level comparisons (Tobyne et al., 2017). Each data set contained around 1 hour of signals represented as 4800 time points, and each subject contained 60,000 vertices. I parcellated each subject’s brain according to the atlas mentioned above, so that each parcel contained the mean time series from its vertices, and correlated the mean time series among all parcels (Pearson). This produced a weighted adjacency matrix for each subject, where each of the 360 parcels is a vertex, and edges the correlations among all of them. Next, I took the exponential of the correlations, Fisher transformed them (i.e. tanh) so that all weights were positive while maintaining the shape of the original correlation distribution, and retained the edges whose adjusted p-values remained significant (FDR < 0.05). Again, I quantified each parcel’s strength and betweenness, averaged them across subjects, and examined whether any of the ROIs showed uniquely high centralities. For this, I performed a permutation comparing the mean strength between the ROIs and the rest of the brain for each participant.

For the main goal of this section, I maximized the strength-based modularity metric as a community detection method (fast-greedy algorithm), so that I could check if all the DMN ROIs from the previous section were affiliated with the same community. While this measure has resolution problems at large and low scales (Fortunato & Hric, 2016), if it does extract a close match to standard DMN networks, it could act as a good method for the final stage of the project (but see Peel, Larremore, and Clauset 2017 for issues with ground-truth). Even though this analysis is mainly geared to confirm that these ROIs are jointly captured, this step also let me examine whether some of the ROIs could be subdivided into DMN and non-DMN regions at this level of granularity.

In order to evaluate the validity of the modularity-based communities, I also clustered the brain by means of spectral partitioning. Specifically, I took the Fiedler vector (eigenvector associated with the second lowest eigenvalue) from each participant’s normalized Laplacian matrix, and divided the brain by the sign of the eigenvector values. This analysis served the additional purpose of more strictly dividing the brain into DMN and non-DMN communities. Importantly, given the high density of these networks (see results), spectral partitioning was unlikely to face the issues associated with its use in sparse networks (Fortunato & Hric, 2016). Finally, to quantify the agreement of these community methods, I computed both the adjusted rand index and variation of information distance (normalized by the logarithm of the number of vertices, 360 in this case) between them per subject. The adjusted rand index denotes the proportion of vertices that coincide in affiliation across partitions, and compares this score to a baseline given by the expectation based on a random vertex assignment for an equal number of clusters across partitions. On the other hand, variation of information gives a general sense of the amount of information to be gained by the complementary community partitioning (Meila, 2007), measured by the addition of the conditional entropies per combination of clusters across partitions. While the rand index is more easily interpreted, it suffers from the assumption that the baseline should have an equal number of vertices per cluster across partitions, which is not necessarily the case. Therefore, I used variation of information as a more robust similarity metric (Fortunato & Hric, 2016). Importantly, due to having just 3 subjects, I abstained from performing formal statistics on the descriptive measures that I show.

**Within-ROI Connectivity**
For the third stage, I “zoomed in” each subject’s brain by computing the correlation among all of the surface-vertices contained within each ROI across all the ROIs, and applied the same correction as in step two. That means that the vertices became smaller functional brain units that are connected by the same type of edge (i.e. adjusted correlations) as in step two. This time, I only focused on partitioning the brain through modularity and spectral partitioning to see if the ROIs can be subdivided at this finer level of resolution. It is worth noting that while communities across subjects should be similar, the idiosyncrasies explained previously (i.e. cortical folding) should produce noticeable differences as well.

This time I added two new areas to the ROIs: 7m, as this area had unique relevance in DMN when compared to valuation (evidenced by its connectivity strength in DMN and chi- squared permutation in the meta-analysis, see results), and V1, as this visual processing area should not be affiliated with DMN. These regions thus acted as dual reference points, such that the community affiliated with 7m should denote DMN, while V1 indicated non-DMN. As before, I tested the validity of the partitions using the adjusted rand index and normalized variation of information. However, this time I also used the rand index to quantify the community correspondence between subjects. Specifically, I divided each subject’s spectrally- partitioned brain into PCC and mPFC topographical zones (posterior and frontal brain, respectively), and computed the index of each region between subjects. The high cross-subject heterogeneity of cortical folding of mPFC should produce relatively low rand indices, while the more homogeneous PCC should display the opposite pattern. This analysis was meant to tackle the subject-specific limitations discussed above. The final patterns were inspected visually.

References {.unnumbered}
========================================

Figures and Tables should be labelled and referenced in the standard way
using the `\label{}` and `\ref{}` commands.

Figure \[fig:frog\] shows an example of how to insert a column-wide
figure. To insert a figure wider than one column, please use the
`\begin{figure*}...\end{figure*}` environment. Figures wider than one
column should be sized to 11.4 cm or 17.8 cm wide.

Single column equations {#single-column-equations .unnumbered}
-----------------------

Authors may use 1- or 2-column equations in their article, according to
their preference.

To allow an equation to span both columns, options are to use the
`\begin{figure*}...\end{figure*}` environment mentioned above for
figures, or to use the `\begin{widetext}...\end{widetext}` environment
as shown in equation \[eqn:example\] below.

Please note that this option may run into problems with floats and
footnotes, as mentioned in the [cuted package
documentation](http://texdoc.net/pkg/cuted). In the case of problems
with footnotes, it may be possible to correct the situation using
commands `\footnotemark` and `\footnotetext`.

$$\begin{aligned}
(x+y)^3&=(x+y)(x+y)^2\\
       &=(x+y)(x^2+2xy+y^2) \label{eqn:example} \\
       &=x^3+3x^2y+3xy^3+x^3. 
\end{aligned}$$

<!-- pandoc writes all tables using longtable, which fails in 2-column mode

  Species                    CBS     CV     G3
  ----------------------- ------ ------ ------
  1\. Acetaldehyde           0.0    0.0    0.0
  2\. Vinyl alcohol          9.1    9.6   13.5
  3\. Hydroxyethylidene     50.8   51.2   54.0

  : Comparison of the fitted potential energy surfaces and ab initio
  benchmark electronic energy calculations

-->

Supporting Information (SI) {#supporting-information-si .unnumbered}
---------------------------

The main text of the paper must stand on its own without the SI. Refer
to SI in the manuscript at an appropriate point in the text. Number
supporting figures and tables starting with S1, S2, etc. Authors are
limited to no more than 10 SI files, not including movie files. Authors
who place detailed materials and methods in SI must provide sufficient
detail in the main text methods to enable a reader to follow the logic
of the procedures and results and also must reference the online
methods. If a paper is fundamentally a study of a new method or
technique, then the methods must be described completely in the main
text. Because PNAS edits SI and composes it into a single PDF, authors
must provide the following file formats only.

### SI Text {#si-text .unnumbered}

Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a
PDF with the same file name for visual reference).

### SI Figures {#si-figures .unnumbered}

Provide a brief legend for each supporting figure after the supporting
text. Provide figure images in TIFF, EPS, high-resolution PDF, JPEG, or
GIF format; figures may not be embedded in manuscript text. When saving
TIFF files, use only LZW compression; do not use JPEG compression. Do
not save figure numbers, legends, or author names as part of the image.
Composite figures must be pre-assembled.

### 3D Figures {#d-figures .unnumbered}

Supply a composable U3D or PRC file so that it may be edited and
composed. Authors may submit a PDF file but please note it will be
published in raw format and will not be edited or composed.

### SI Tables {#si-tables .unnumbered}

Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a
PDF with the same file name for visual reference); include only one
table per file. Do not use tabs or spaces to separate columns in Word
tables.

### SI Datasets {#si-datasets .unnumbered}

Supply Excel (.xls), RTF, or PDF files. This file type will be published
in raw format and will not be edited or composed.

### SI Movies {#si-movies .unnumbered}

Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media
(wmv), animated GIF (gif), or MPEG files and submit a brief legend for
each movie in a Word or RTF file. All movies should be submitted at the
desired reproduction size and length. Movies should be no more than 10
MB in size.

### Still images {#still-images .unnumbered}

Authors must provide a still image from each video file. Supply TIFF,
EPS, high-resolution PDF, JPEG, or GIF files.

### Appendices {#appendices .unnumbered}

PNAS prefers that authors submit individual source files to ensure
readability. If this is not possible, supply a single PDF file that
contains all of the SI associated with the paper. This file type will be
published in raw format and will not be edited or composed.

<!-- Leave these lines as they are at the end of your .Rmd file to ensure placement of methods & acknowledgements sections before the references-->
\showmatmethods
\showacknow
\pnasbreak